### 2015-02-14 rmd/paper.Rmd
cutting with great abandon

---
title: 'Claret: Using Data Types for Highly Concurrent Distributed Transactions'
preprint: true

author:
  - {family: Holt,  given: Brandon, affiliation: 1, email: bholt}
  - {family: Zhang, given: Irene,   affiliation: 1, email: iyzhang}
  - {family: Ports, given: Dan,     affiliation: 1, email: dkp}
  - {family: Oskin, given: Mark,    affiliation: 1, email: oskin}
  - {family: Ceze,  given: Luis,    affiliation: 1, email: luisceze}

organization:
  - {id: 1, name: University of Washington}

conference:
  name: PaPoC 2015
  location: 'April 21, 2015, Bordeaux, France'
  year: 2015

doi: 0

layout: sigplanconf
bibliography: biblio.bib
output:
  pdf_document:
    fig_caption: yes

abstract: |
    Out of the many NoSQL databases in use today, some that provide simple data structures for records, such as Redis and MongoDB, are now becoming popular. Building applications using these data structures rather than plain string values provides programmers with a way to communicate intent to the database system without sacrificing flexibility or committing to a fixed schema. Currently, this expressiveness is used to ensure related values are co-located so they can be quickly accessed together (e.g. maps and other aggregates) and to provide complex atomic operations (e.g. set insertion). However, there are many more ways in which data types can be used to make databases more efficient and simpler to use that are not yet being exploited.

    In this work, we demonstrate several ways of leveraging data structure semantics in databases, focusing primarily on commutativity. Reasoning about operation reordering can allow transactions to execute concurrently that would conflict under traditional concurrency control. Using Retwis, a Twitter clone built for Redis, as a case study, we show that using commutativity can reduce transaction abort rates for high-contention, update-heavy workloads that arise in real social networks. We conclude that data types are a good abstraction for database records, providing a safe and expressive programming model with ample opportunities for optimization, which will make databases more safe and scalable.
---
```{r setup, include=FALSE}
opts_chunk$set(dev='pdf', echo=F, message=F, warning=F, error=F, fig.width=3.2, fig.height=3)
```

# Introduction

The move to non-relational (NoSQL) databases was motivated by a desire for scalability and flexibility. People found that by not enforcing strong consistency, they could better scale services to millions or billions of users while meeting tight performance goals. By forgoing the structure of relational schemas, they gained more direct control over performance at the cost of significant automatic optimization.
The observation has been that relaxed or eventual consistency is a good fit for many applications built on top of these NoSQL because potential conflicts are difficult to observe in practice and that users are likely to accept minor inconsistencies, such as two tweets being out of temporal order on their Twitter timeline. However, this leaves much to chance; there is likely no guarantee that more significant inconsistencies are impossible.

In situations where consistency really is crucial, developers can enforce stronger guarantees on top of any platform, but they often get it wrong. Google, recognizing the benefit to programmer productivity, reintroduced serializable transactions to their next generation database, Spanner \cite{Spanner}. However, this leaves them with two extreme choices, with significant performance impact. If certain parts of an application can tolerate imprecision, then why not capture those properties in the database programming model? What if the programmer could express the semantics they desire succinctly and precisely, in a way that the database can better reason about conflicts and optimize performance, without giving up the flexibility and scalability of their existing systems?

Abstract data types (ADTs) are the solution to this problem. Rather than limiting the records in databases to being primitive types like strings or integers, raising them to more complex data types provides a richer interface, exposing ample opportunities for optimization to the database and a precise mechanism to express the intentions of programmers. Understanding the properties of ADT operations, such as which operations commute, can be used in a variety of ways to improve database performance, such as reducing the number of aborts. By choosing the right data type, programmers make tradeoffs between performance and precision based on their application's needs.

# Commutativity {#comm}

Commutativity is well known, especially in distributed systems, for enabling important optimizations. Since the 80s, commutativity has been exploited by database systems designers \cite{Weihl:1988,Fekete:90}, within the safe confines of relational models, where knowledge of query plans and complete control of the data structures allows systems to determine when transactions may conflict. Recently, commutativity has seen a resurgence in systems without a predefined data model, such as NoSQL databases and transactional memory.
Commutativity is leveraged in eventually consistent databases for convergence in work such as RedBlue consistency \cite{Li:OSDI12} and conflict-free replicated data types (CRDTs) \cite{Shapiro:SSS11}. Other systems specialize for commutative operations to improve transaction processing, such as Lynx \cite{Zhang:SOSP13} for tracking serializability, and Doppel \cite{Narula:OSDI14} for executing operations in parallel on highly contended records, and HyFlow \cite{Kim:EuroPar} for reordering operations in the context of distributed transactional memory. We propose unifying and generalizing these uses of commutativity under the abstraction afforded by ADTs.

## Commutativity Specifications

\begin{table}
\resizebox{\columnwidth}{!}{%
\centering
\begin{tabular}{lll}
\textbf{method:} & \textbf{commutes with:} & \textbf{when:} \\
\hline
\texttt{add(x): void} & \texttt{add(y)} & $\forall x, y$ \\
\texttt{remove(x): void} & \texttt{remove(y)} & $\forall x, y$ \\
    & \texttt{add(y)} & $x \ne y$ \\
\texttt{size(): int} & \texttt{add(x)} & $x \in Set$ \\
    & \texttt{remove(x)} & $x \notin Set$ \\
\texttt{contains(x): bool} & \texttt{add(y)} & $x \ne y \lor y \in Set$ \\
    & \texttt{remove(y)} & $x \ne y \lor y \notin Set$ \\
    & \texttt{size()} & $\forall x$ \\
    \hline
\end{tabular}
\caption{\label{spec} Commutativity Specification for Set.}}
\end{table}

Though *commutativity* is often discussed in terms of an operation commuting with all other operations, it is actually more nuanced. If a pair of operations commute, then executing them in either order will produce the same result. Using the definitions from \cite{Kulkarni:PLDI11}, whether or not a pair of method invocations commute is a function of the methods, their arguments, their return values, and the *abstract* state of their target. We call the full set of commutativity rules for an ADT its *commutativity specification.* An example specification for a *Set* is shown in \autoref{spec}. It is important to note that interface choice affects commutativity: if `add(x)` returned a boolean that expressed if the item was added or not, then `add(x)` would only commute with `add(y)` if $x \ne y$. In this case, the difference doesn't drastically impact commutativity, but it would affect the cost of checking commutativity dynamically. There are actually many valid specifications which expose less than the maximum commutativity, but may be cheaper to implement.\cite{Kulkarni:PLDI11} ADT designers can navigate these choices to optimize an implementation.

## Transactional Boosting
If two operations on the same record in two different transactions commute, then the transactions can safely execute concurrently, even though they both update the record. This technique of raising the level of abstraction in transactions to operations is known as *transactional boosting* \cite{Herlihy:PPoPP08}. This straightforward use of commutativity was shown to significantly improve abort rates in state-of-the-art software transactional memory. In \autoref{eval}, we show how it can be applied to distributed transactions.

## Other opportunities

Sometimes, due to heavily skewed workloads such as those arising from social networks, there may be so many requests to a single record that just executing all the operations sequentially is a bottleneck which prevents scaling. In addition to transactional boosting, which we evaluate in this work, there are many other uses for commutativity. We mention just a few here which we have yet to evaluate.

**Record splitting.**
Associativity, often paired with commutativity, can be leveraged to execute operations in parallel on a "split" record and the changes made to these splits can be merged to form a consistent whole. Doppel \cite{Narula:OSDI14} used this for a handful of operations. We propose generalizing this concept to ADTs: associative operations execute on replicas first, other operations wait to execute until the replicas have been brought back in sync.

**Combining.**
Another spin on associative operations is to merge or *combine* operations as they come in. This is known as combining \cite{flatCombining,yew:combining-trees,funnels}, and it can drastically reduce contention. Combining can be done hierarchically: first with a few neighbors, then with clusters of neighbors, finally the combined operation is applied to the shared data structure.

# Evaluation {#eval}
To demonstrate the efficacy of leveraging commutative operations in transactions, we built a simple prototype key-value store, modeled after Redis, that supports complex data types for records, each with their own set of operations. 
The transaction protocol employs standard two-phase commit, and two-phase locking; more details are in the appendix (\autoref{apx:protocol}).

Commutativity comes into play in the locking scheme. Using the algorithms from \cite{Kulkarni:PLDI11} and our commutativity specifications, we design an abstract lock for each record type. Our `SortedSet`, for instance, has an `add` mode which allows all insertion operations to commute, but disallows operations like `contains` or `size`.
As a baseline, we implement a standard reader/writer locking scheme that allows all read-only operations to execute concurrently, but enforces that only one transaction may modify a record at a time.

These experiments were run with 4 shards on 4 local nodes, each with 8-core 2GHz Xeon E5335 processors.

```{r stress, include=F}
d <- data(db("
    select * from tapir where 
    total_time is not null
    and name like 'stress-v0.14%'
    ", factors=c('nshards', 'nclients'), numeric=c('total_time', 'txn_count')))
d$opmix <- factor(revalue(d$mix, c(
    'mostly_update'='35% read\n65% update',
    'update_heavy'='50% read\n50% update',
    'read_heavy'='90% read\n10% update'
)))
d$dist <- factor(revalue(d$alpha, c('0.6'='Zipf: 0.6', '-1'='Uniform')))
d.u <- subset(d, nshards == 4 & nkeys == 10000 & (alpha == '0.6' | alpha == '-1') & grepl('update_heavy|read_heavy', mix))
ggplot(d.u, aes(x=nclients, y=throughput/1000, group=cc, fill=cc, color=cc, linetype=cc))+
    stat_summary(fun.y=max, geom="line")+
    xlab('Concurrent clients')+ylab('Throughput (k/sec)')+
    expand_limits(y=0)+
    facet_grid(dist~opmix)+
    theme_mine+theme(legend.position='bottom', legend.direction='horizontal', legend.title.align=1)+
    cc_scales(title='Concurrency\ncontrol:')
```
\begin{figure}[t]
\centering
\includegraphics{figure/stress-1.pdf}
\caption{Throughput of raw Set operations.\label{stress_tput}}
\end{figure}

**Microbenchmark: Set operations.** We first evaluate performance with a simple workload consisting of a raw mix of `Set` operations randomly distributed over 10,000 keys. We use both a uniform random distribution as well as a skewed Zipfian distribution with a coefficient of 0.6. In \autoref{stress_tput}, we see that commutative transactions perform strictly better, showing the most pronounced benefit over the more update-heavy, skewed workload.

```{r throughput, include=F}
d <- data(db("
    select * from tapir where 
    generator_time is not null and total_time is not null
    and name like 'claret-v0.14%'
  ",
    factors=c('nshards', 'nclients'),
    numeric=c('total_time', 'txn_count')
))
d.u <- subset(d, nshards == 4 & initusers == 4096 & grepl('geom_repost|read_heavy', mix))
ggplot(d.u, aes(x=nclients, y=throughput, group=cc, fill=cc, color=cc, linetype=cc))+
    stat_summary(fun.y=mean, geom="line")+
    xlab('Concurrent clients')+ylab('Throughput (transactions / sec)')+
    expand_limits(y=0)+
    facet_wrap(~workload)+
    theme_mine+theme(legend.position='bottom', legend.direction='horizontal', legend.title.align=1)+
    cc_scales(title='Concurrency\ncontrol:')
```
\begin{figure}[t]
\centering
\includegraphics{figure/throughput-1.pdf}
\caption{Throughput of social network workload (Retwis) with 4000 users.
Leveraging commutativity prevents performance from falling over even
when posts spread virally (repost-heavy).\label{tput}}
\end{figure}

**Case study: Retwis.** 
To understand performance of more typical of web workloads, we implemented a version of *Retwis*, a simplified Twitter clone designed originally for Redis. We use data structures such as sets to track each user's followers and posts, and keep a materialized up-to-date timeline for each user (represented as a sorted set). On top of Retwis's basic functionality, we also add a "repost" option that works like Twitter's "retweet". We simulate a realistic workload using a synthetic graph with power-law degree distribution and clients that randomly select between Retwis transactions including "add follower", "new post", and "repost", executing them as fast as they can. More about this workload is in the appendix (\autoref{apx:retwis}).

\autoref{tput} shows the results of this simulation. When most of the traffic is content consumption (reading timelines), both systems perform well enough. However, when we simulate a workload where clients repost popular posts from their timelines, we see a viral propagation effect, where a large fraction of the users get and share a post. As Twitter came to a standstill when Ellen Degeneres's Oscar selfie set a retweeting record, so too does our baseline fall over. But with commutativity, we see that performance continues to scale even under this arduous load.

# Appendix

## Transaction protocol {#apx:protocol}

Our protocol for executing transactions is fairly straightforward: two-phase commit, and uses two-phase locking with retries to guarantee isolation. We employ a number of standard optimizations, such as delaying acquiring locks for operations that don't return a value to the *prepare* step so that locks are held for as short a time as possible. However, there is one step that is non-standard in order to support complex data types where rolling back state changes would be non-trivial.

To support transactions with arbitrary data structure operations, each operation is split into two steps: *stage* and *apply*. During transaction execution, each operation's *stage* method attempts to acquire the necessary lock and may return a value *as if the operation has completed* (e.g. an "increment" speculatively returns the incremented value). When the transaction is prepared to commit, *apply* is called on each staged operation to actually mutate the underlying data structure. This allows operations to easily be un-staged if the transaction fails to acquire all the necessary locks, without requiring rollbacks.

## Retwis workload {#apx:retwis}
```{r}
df <- subset(db("select * from tapir where stat_following_counts is not null and name like '%v0.14%'"),
  nclients == 32
  & initusers == 4096
)
df$grp <- with(df, sprintf("%s\n%s\nmix:%s/%s,\n%s", name, ccmode, mix, alpha, gen))

histogram.facets <- function(df, measure, grp) {
  d <- data.frame(x=c(),y=c(),version=c())
  for (i in 1:nrow(df)) {
    d <- rbind(d, df.histogram(df[i,measure], df[i,grp]))
  }
  return(d)
}
```

```{r followers, include=F}
d.follow <- histogram.facets(subset(df, initusers == 4096 & mix == 'geom_repost'), 'stat_follower_counts', 'grp')
ggplot(d.follow, aes(x=x, weight=y))+
    stat_ecdf(color=c.blue)+
    xlab('# followers / user (log scale)')+ylab('CDF (log scale)')+
    scale_x_log10(breaks=c(1,10,100,1000))+scale_y_log10(breaks=c(0.1,0.5,1.0))+
    theme_mine
```
\begin{figure}[t]\centering
\includegraphics{figure/followers-1.pdf}
\caption{CDF of the number of followers for users generated by the
Kronecker synthetic graph generator, matching the power-law degree
distribution of natural graphs.\label{followers}}
\end{figure}

```{r reposts, include=F}
d.repost <- histogram.facets(subset(df, initusers == 4096 & mix == 'geom_repost'), 'stat_repost_counts', 'grp')
ggplot(d.repost, aes(x=x, weight=y))+
    stat_ecdf(color=c.blue)+
    xlab('# reposts')+ylab('count')+
    scale_x_log10(breaks=c(1,10,100,1000))+scale_y_log10(breaks=c(0.1,0.2,0.4,0.6,0.8,1.0))+
    xlab('# reposts (log scale)')+ylab('CDF (log scale)')+
    theme_mine
```
\begin{figure}[t]\centering
\includegraphics{figure/reposts-1.pdf}
\caption{CDF of the number of times a post is reposted, matching traces
from real workloads. Due to the power-law graph structure, if users tend
to reposts popular recent posts on their timeline, it results in another
power-law distribution. Note that that some posts are reposted to over a
quarter of the graph (4000 total users).\label{reposts}}
\end{figure}

With our Retwis workload, rather than approximating the skew in real-world workloads with a Zipfian as many other systems do, we attempt to simulate the behavior of social networks with a realistic synthetic graph and a simple model of user behavior for posting and reposting.

For our synthetic graph, we use the Kronecker graph generator from the Graph 500 benchmark \cite{graph500}. This generator is designed to result in graphs with the same power-law degree distribution found in natural graphs. \autoref{followers} shows the cumulative distribution function (CDF) of the number of followers per user for the synthetic graph of approximately 4000 users, with an average number of followers of 16 (scale 12 with edgefactor of 16 in Graph500's terms) which we use in our performance results back in \autoref{eval}. Most users should have relatively few followers; we see that roughly 50% have fewer than 100 followers, while a very small number of users have over 1000 followers.

We use a simple model of user behavior to determine when and which posts to repost. Each time we load the most recent posts in a timeline for a random user (uniformly selected), they are sorted by the number of times they have already been reposted, and a discrete geometric distribution, skewed toward 0, is used to select the number of these to repost. This results in the "viral" propagation effect that is observed in real social networks. \autoref{reposts} shows the distribution of the number of times a post was reposted, which is again a power-law distribution. Note that a small number of posts are reposted so much that they end up on over a quarter of users' timelines.